---
title: "N and P relationships - visual exploration"
author: "Kim Cressman"
date: "`r Sys.Date()`"
output: 
    html_document:
        keep_md: true
        code_folding: 'hide'
        fig_width: 8
        fig_height: 4
---
```{r, message=FALSE, warning=FALSE}
library(tidyverse)
library(lubridate)

setwd("C:/Users/kimberly.cressman/Desktop/Main Docs/Phosphate Publication/gndphosphate/Kim")
load("../data/nut_dat.RData")
```

### Data processing code chunk  
Expand this if you want to check either my reshaping of the dataset (I loaded `nut_dat.RData` from the gndphosphate GitHub repo) or my math going from mg/L to micromolar. The math I used was: 

+  PuM = Pmg/L * 1000 / 30.97  
+  NuM = Nmg/L * 1000 / 14  
+  DIN: NO23 + NH4, after converting both forms of N to micromolar


```{r, warning=FALSE}
BL <- nut_dat %>%
    filter(StationCode=="BL")

# make it wide; add in columns for molarity (MICROMOLAR is the unit) and ratios
BL <- BL %>%
    select(TimeFrame, date, nutrient, value) %>%
    spread(key=nutrient, value=value) %>%
    mutate(PO4M = PO4F * 1000/30.97, 
           NO23M = NO23F * 1000/14, 
           NH4M = NH4F * 1000/14,
           DINM = NH4M + NO23M,
           NtoP = DINM/PO4M,
           NH4toP = NH4M/PO4M)
```

******

### Scatterplot of DIN (as NO23 + NH4) and PO4, colored by time period 

+  These have been converted to micromolar.  
+  I ignored "below detection" data points and just allowed 0.01 to be used.

```{r, warning=FALSE}
# scatter plot of DIN and PO4, colored by time
ggplot(BL) +
    geom_point(aes(x=DINM, y=PO4M, col=TimeFrame), size=3) +
    scale_y_log10() +
    scale_x_log10() +
    theme_bw() +
    ggtitle("Scatterplot of PO4 and DIN across time") +
    xlab("DIN (= NO23 + NH4; uM)") +
    ylab("PO4 (uM)")
```

******

### Time series of DIN (as NO23 + NH4) to PO4 ratio  

+  These were converted to micromolar.  
+  I again ignored "below detection" data points. 

```{r, warning=FALSE}
# time series, as points, of N:P molar ratio
ggplot(BL) +
    geom_point(aes(x=date, y=NtoP, col=TimeFrame), size=3) +
    theme_bw() +
    ggtitle("Molar N:P ratio through time") +
    xlab("Date") +
    ylab("DIN (as NO23 + NH4):P")
```

******

### Time series of NH4:PO4 ratio 

+  Made this one because NO23 is almost always below detection, and NH4 almost always _isn't_.  
+  Again, converted to micromolar and ignoring "below detection" data points.  
+  This looks almost exactly like the DIN:P ratio graph, with a few exceptions in NI1.

```{r, warning=FALSE}
# and NH4:P
ggplot(BL) +
    geom_point(aes(x=date, y=NH4toP, col=TimeFrame), size=3) +
    theme_bw() +
    ggtitle("Molar NH4:P ratio through time") +
    xlab("Date") +
    ylab("NH4:P")
```


******

__Housekeeping__  

```{r}
sessionInfo()
Sys.info()
```

