---
title: "Nutrients below detection"
author: "Kim Cressman"
date: "`r Sys.Date()`"
output: 
    html_document:
        keep_md: true
        code_folding: 'hide'
        fig_width: 8
        fig_height: 4
---
```{r, message=FALSE, warning=FALSE}
library(tidyverse)
library(lubridate)

setwd("C:/Users/kimberly.cressman/Desktop/Main Docs/Phosphate Publication/gndphosphate/Kim")
load("../data/nut_dat.RData")
```

### Data processing code chunk  
Expand this if you want to check either my reshaping of the dataset (I loaded `nut_dat.RData` from the gndphosphate GitHub repo) or my math going from mg/L to micromolar. The math I used was: 

+  PuM = Pmg/L * 1000 / 30.97  
+  NuM = Nmg/L * 1000 / 14  
+  DIN: NO23 + NH4, after converting both forms of N to micromolar


```{r, warning=FALSE}
# make it wide; add in columns for molarity (MICROMOLAR is the unit) and ratios
nut_wide <- nut_dat %>%
    select(StationCode, TimeFrame, date, nutrient, value) %>%
    spread(key=nutrient, value=value) %>%
    mutate(PO4M = PO4F * 1000/30.97, 
           NO23M = NO23F * 1000/14, 
           NH4M = NH4F * 1000/14,
           DINM = NH4M + NO23M,
           NtoP = DINM/PO4M,
           NH4toP = NH4M/PO4M)
```


The nutrient values have already been floored at 0.01, the highest MDL. Exactly how many 0.01s are there for each nutrient, at each site, in each time period?  

Note that some 0.01s could be actual above-detection 0.01 values, and I'm ignoring that fact here. It's fairly minor compared to the amount of data (admittedly, this is an assumption that should probably be chedked down the line).  

Actually, that long data format should work better for answering this question.


```{r}
nut_bd <- nut_dat %>%
    group_by(StationCode, nutrient, TimeFrame) %>%
    summarize(bdvalues = sum(value == 0.01, na.rm=TRUE), 
              totalvalues = length(value)) %>%
    mutate(prop_bd = round(bdvalues/totalvalues, 3))

```

What I want to do now is make either one nice pretty table, or separate tables for each nutrient; where rows are the site and columns are the time period, and values are the proporation of below detection. I don't know off the top of my head how to do it, and it will have to wait until after some lab work.  

`split` is easy, and then it should just be a `spread` command.... i think.  


Trying to use knitr functions to print the whole darn data frame. This isn't pretty, but it's all there.

```{r}
library(knitr)
kable(nut_bd, format="html")
```



******

_Housekeeping_

```{r}
sessionInfo()
```

